---
title: "AE1812 Zim RDA"
author: "Setta"
date: "August 5, 2020"
output: html_document
---

#### Load in data and libraries

```{r setup, include=FALSE}
setwd("WorkingDir/AE1812_MiSeq_11182019/Cutadapt_primersRemoved/zim")
```

# Load packages
```{r, include=FALSE}
# BiocManager::install("DESeq2")
library("DESeq2")
library("phyloseq")
library("vegan")
library("ggplot2")
library("tidyr")
library("viridis")
library("tidyverse")
library("RColorBrewer")
library("dendextend")
##  Install packages for Gloor et al. 2017:
library("knitr")
# BiocManager::install("ALDEx2")
library("ALDEx2")
# install.packages('devtools')
# devtools::install_github('ggloor/CoDaSeq/CoDaSeq')
library("CoDaSeq")
library("zCompositions")
library("igraph")
library("grDevices")
library("propr")
library("car")
```

### Load data
```{r, results='hide'}
# Count Table:
zcount_tab <- read.csv("Zim_BacInsitu200Filt_CountTab.csv", header=T, row.names=1,check.names=F, sep=",")
# Taxonomy File:
ztax_tab_pr2 <- as.matrix(read.csv("Zim_BacInsitu200Filt_TaxTab.csv", header=T, row.names=1,check.names=F, sep=","))
# Info Table:
zsample_info_tab <- read.csv("Zim_BacInsitu200Filt_InfoTab.csv", header=T, row.names=1,check.names=F, sep=",")
# Look into sample info table:
str(zsample_info_tab)
# read in pore size as a factor:
zsample_info_tab$Pore_Size_um<-as.factor(zsample_info_tab$Pore_Size_um)
# read in DCM as numerical
zsample_info_tab$DCM_m<-as.numeric(zsample_info_tab$DCM_m)
# read in Region as a factor
zsample_info_tab$Region<-as.factor(zsample_info_tab$Region)
# read in Sample_or_Blank as a factor
zsample_info_tab$Sample_or_Blank<-as.factor(zsample_info_tab$Sample_or_Blank)
# read in Exp_Region as a factor
zsample_info_tab$Exp_Region<-as.factor(zsample_info_tab$Exp_Region)
# read in CastID as a factor
zsample_info_tab$CastID<-as.factor(zsample_info_tab$CastID)
# read in InSitu_Exp as a factor
zsample_info_tab$InSitu_Exp<-as.factor(zsample_info_tab$InSitu_Exp)
# read in Treatment as a factor
zsample_info_tab$Treatment<-as.factor(zsample_info_tab$Treatment)
# check data structure again:
str(zsample_info_tab)

# Create phyloseq object using our un-transformed count table
zcount_tab_phy <- otu_table(zcount_tab, taxa_are_rows=T)
ztax_tab_pr2_phy <- tax_table(ztax_tab_pr2)
zsample_info_tab_phy <- sample_data(zsample_info_tab)
z_ASV_physeq <- phyloseq(zcount_tab_phy, ztax_tab_pr2_phy, zsample_info_tab_phy)

```

### Remove samples without environmental variables
```{r}
# Cast 52 (SPS 34 & 35, S20A) does not have any environmental variables
z_insitu<-subset_samples(z_insitu, Cast!="52")

# remove samples without complete environmental variables:
z_insitu<-subset_samples(z_insitu, SampleID!="AE1812 197" & SampleID!="AE1812 336" & SampleID!="AE1812 340")
z_insitu_otu<-t(otu_table(z_insitu))
```

## Part One: Transformations of raw data

### Step 1: Hellinger & Cord Transformation of data using vegan
```{r}
# remove samples without complete environmental variables:
z_insitu<-subset_samples(z_insitu, SampleID!="AE1812 197" & SampleID!="AE1812 336" & SampleID!="AE1812 340")
z_insitu_otu<-t(otu_table(z_insitu))
# Incubations
z_incub_otu<-t(otu_table(z_incub))
# Transform data using the vegan function decostand (standardization methods for community ecology)
# first for the cord transformation:
z_insitu_cord<-decostand(z_insitu_otu, method="normalize")
z_incub_cord<-decostand(z_incub_otu, method="normalize")
# next for the hellinger transformation:
z_insitu_hell<-decostand(z_insitu_otu, method="hellinger")
z_incub_hell<-decostand(z_incub_otu, method="hellinger")
```

### Step 2: Transformation of Environmental Data

Check to see if data is normal to begin, and if there are any outliers in data etc.
```{r}
# Environmental components:
z_insitu_env<-data.frame(sample_data(z_insitu))
# First check and see if environmental data is normal - Insitu env variables
z_insitu_env
qqPlot(z_insitu_env$SiO2_umolL)
qqPlot(z_insitu_env$Nit_Nit_umolL)
qqPlot(z_insitu_env$DON_umolL)
qqPlot(z_insitu_env$SRP_umolL)
qqPlot(z_insitu_env$temperature)
qqPlot(z_insitu_env$CPAR)
# most are not normal

# most are not normal
z_insitu_env1<-z_insitu_env[,c(3,4,8,9,26,30)]

z_insitu_env1<-z_insitu_env1[complete.cases(z_insitu_env1),]

pairs(z_insitu_env1, lower.panel = NULL)
# DIN and Nit_Nit correlated...can't do both, we'll use Nit_Nit because data set is more full (Amm only above detection in some samples)
```

After checking that there are outliers and most data is not normally transformed we can apply a  pseudocount (+1) and log10 transformation
```{r}
# Transform M3 samples with all variables first
z_insitu_env_transf<-log10(z_insitu_env1)
```

Look at normality in data again:
```{r}
# In Situ Samples First
z_insitu_env_transf
qqPlot(z_insitu_env_transf$SiO2_umolL)
qqPlot(z_insitu_env_transf$Nit_Nit_umolL)
qqPlot(z_insitu_env_transf$DON_umolL)
qqPlot(z_insitu_env_transf$SRP_umolL)
qqPlot(z_insitu_env_transf$temperature)
qqPlot(z_insitu_env_transf$CPAR)

# most look better - will find out if these transformations make a difference in RDAs
```


I will keep the log transformation of Nit_Nit, DON and CPAR:
```{r}
# In Situ Samples First
z_insitu_env2<-z_insitu_env1
# Nit_Nit transformation:
z_insitu_env2$Log_Nit<-log10(z_insitu_env2$Nit_Nit_umolL)
# SRP transformation:
z_insitu_env2$Log_SRP<-log10(z_insitu_env2$SRP_umolL)
# CPAR transformation:
z_insitu_env2$Log_CPAR<-log10(z_insitu_env2$CPAR)

# All variables to use for RDA
zim_env_rda<-z_insitu_env2[,c(1,3,5,7:9)]
```


## Part Two: RDA of all combinations of transformed and untransformed data

Now we have hellinger transformed data and a few environmental variables log transformed (Nit_Nit, SRP, CPAR).

Will create an RDA and PERMANOVA with these to see if significant associations change based on transformations.

##### Hellinger Transformed

In situ samples original data with no environmental transformation
```{r}
# First run RDA
insitu_rda_tree = rda(z_insitu_otu ~ SiO2_umolL + Nit_Nit_umolL + DON_umolL + SRP_umolL + temperature + CPAR, data=z_insitu_env1)
insitu_rda_tree
# Calculate R-squared value
RsquareAdj(insitu_rda_tree)
# Look at Samples that are driven by environmental variables
par(mfrow=c(1, 1))
plot(insitu_rda_tree, type='n', scaling=1)
orditorp(insitu_rda_tree, display='site', cex=0.9, scaling=1, col='blue')
text(insitu_rda_tree, display='cn', col='red')
# Look at OTUs that are driven by environmental variables:
plot(insitu_rda_tree, type='n', scaling=1)
orditorp(insitu_rda_tree, display='sp', cex=0.9, scaling=1, col='black')
text(insitu_rda_tree, display='cn', col='red')
# Run Permanova statistical test
adonis2(z_insitu_otu ~ SiO2_umolL + Nit_Nit_umolL + DON_umolL + SRP_umolL + temperature + CPAR, data=z_insitu_env1)
# Silicate, Nit/Nit, DON, SRP significant

```


In situ samples transformed data with environmental transformation (as needed)
```{r}
# First run RDA
insitu_rda_tree2 = rda(z_insitu_hell ~ SiO2_umolL + DON_umolL + Log_Nit + Log_SRP , data=zim_env_rda)
insitu_rda_tree2
# Calculate R-squared value
RsquareAdj(insitu_rda_tree)
# Look at Samples that are driven by environmental variables
par(mfrow=c(1, 1))
plot(insitu_rda_tree2, type='n', scaling=1)
orditorp(insitu_rda_tree2, display='site', cex=0.9, scaling=1, col='blue')
text(insitu_rda_tree2, display='cn', col='red')
# Look at OTUs that are driven by environmental variables:
plot(insitu_rda_tree2, type='n', scaling=1)
orditorp(insitu_rda_tree2, display='sp', cex=0.9, scaling=1, col='black')
text(insitu_rda_tree2, display='cn', col='red')
# Run Permanova statistical test
adonis2(z_insitu_hell ~ SiO2_umolL + DON_umolL + Log_Nit + Log_SRP , data=zim_env_rda)
# Silicate, Nit/Nit, DON, SRP significant

```

